\subsection{Model Coupling}

Climate scientists create independent simulations that each model a different
part of the planet's ecosystem. To model more complex aspects of the ecosystem,
and for more accurate results, multiple simulations are required to work
together.

There are three ways of doing having models work together \cite{Thevenin}. The
first is merging the model code bases together which generates a single
executable with all models having efficient data exchanges since all data is
contained within the one memory space. This solution isn't ideal since merging
code bases is difficult and further development of the individual models is now
more difficult.

The second is to use a communication library directly, such as MPI. Each model
can keep its own codebase with only explicit communication code being added to
each model. This solution also isn't ideal since the communication code isn't
generic since each model coupling will require its own communication code and
coupling more than two models together results in very complicated exchanges
being implemented.

The third solution is to use a model coupling library which abstracts the low
level communication complexities and acts as a ``middle man'' between each
model. For example, the LES has been coupled to the Weather Research and
Forecasting Model (WRF) using a number of model coupling libraries. This has led
to more accurate wind velocities being predicted compared with the predictions
from WRF alone \cite{Kinbara2010,Nakayama1998}, showing the benefit of having
multiple models interact with one another. There are many model coupling
frameworks available, including: the Model Coupling Toolkit (MCT)
\cite{Jacob2005,Larson2005}; OASIS \cite{Valcke2013,Valcke}; the Earth System
Modeling Framework (ESMF) \cite{Ramework2004}; and OpenPALM
\cite{Piacentini2011}.

\subsection{GMCF}

GMCF is a relatively new framework with development beginning in 2014. GMCF's
primary function is for model coupling. One of long term goals of GMCF is to
automate as much of the model coupling process as possible. Current frameworks
currently require hand modification of existing model code and the writing of
additional code to describe and control the communication between models. This
can present a level of difficulty that a research team may not be willing to
accept so a framework that can automate a significant amount of the model
coupling work would be welcomed.

GMCF also aims to change the way model coupling is organised at runtime. The
current model coupling frameworks are focussed on clusters with single-core
machines and do not take into account the uptake in multicore and heterogeneous
nodes in modern clusters. The focus on single core nodes leads to a
communication that is necessarily distributed using MPI or equivalents and load
balancing between models is organised such that a model gets a fixed number of
nodes, proportional to its runtime compared with the runtimes of the other
models. With the advent of multicore clusters, a new load balancing technique
can be implemented, which is what GMCF does. GMCF aims to limit the coupling of
models to processes within a single node, with limited distributed memory
communication. The idea behind this is that communication in the cluster is more
symmetrical and each node can balance its load more evenly for improved
performance.

To date, GMCF has been used to couple the LES with WRF, acting as a proof of
concept to the new ideas behind the framework.
